---
title: "Lab 11 - Smoking during pregnancy"
author: "Conor Lacey"
date: "03/27/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
suppressWarnings(library(tidyverse)) 
suppressWarnings(library(openintro))
suppressWarnings(library(skimr))
suppressWarnings(library(Hmisc))
suppressWarnings(library(infer))
```

### Exercise 1

```{r ncbirths data}
set.seed(4215)
data("ncbirths")

skim(ncbirths)

ncbirths_num <- ncbirths %>% select_if(is.numeric)
hist.data.frame(ncbirths_num)
```

Variables mature, premie, marital, lowbirthweight, gender, habit, and whitemom are categorical. 

Variables fage, mage, weeks, visits, gained, and weight are numerical. 

It also appears from the histograms that visits and weeks have some outliers.

### Exercise 2

```{r ncbirths_white}
ncbirths_white <- ncbirths %>% filter(whitemom == "white")
mean(ncbirths_white$weight)
```

### Exercise 3 

I believe so. We are wanting to know whether the average weight of Caucasian babies has changed since 1995. However, we only have a sample of data, but we can use this sample to *infer* what is happening in the population. We can also resample from this multiple times and do a bootstap (a simulation) to make this inference.

```{r bootstrap}
avg<-c(NA)
for (i in 1:1000){
x <- sample(ncbirths_white$weight, size = 50, replace = TRUE)
avg[i]<-mean(x)
}
df_weight <- data.frame(avg = avg)
df_weight <- df_weight %>% mutate(avg =(avg + (7.43-mean(ncbirths_white$weight))))

df_weight %>% ggplot(aes(x = avg)) +
  geom_histogram(color = "black",fill = "blue", alpha = 0.4) +
  xlab("weight")

sum((df_weight$avg >= (mean(ncbirths_white$weight))) / length(df_weight$avg)) 
```

.817 is the p-value of the proportion bootstrap samples that yielded a sample mean at least as extreme as the observed sample mean (7.25)
